### 2022-12-09 SALT-abstract/SALT-abstract.tex
SALT abstract final final

\vspace{-.6\baselineskip}
\noindent 
	We present experimental evidence \textbf{(i)} that the projection of the content of attitude complements differs between entailment-canceling operators, and \textbf{(ii)} that this by-operator variation differs across attitude predicates. Our results do not align with the long-standing distinction of factive vs. semi-factive predicates (see, e.g., \citealp{karttunen_observations_1971,hooper_applicability_1973,djarv_cognitive_2018}).
	Instead, the observed by-operator variation reveals a more nuanced picture of lexical semantic/pragma-tic properties which raises important questions for future research on projection.

@@

\vspace{-.4\baselineskip}
	\citet{karttunen_observations_1971} proposed that the CC of factive predicates (e.g. \emph{regret, forget}) projects across all four operators, whereas that of semi-factive predicates (e.g. \emph{discover, realize, see, notice}) always projects from under negation, but not always from under polar questions, modals, or conditionals. 
	To date, systematic investigations of Karttunen's intuition are scarce.

@@

\noindent
{\bf Discussion: Novel research question.}
	Can the observed interaction between predicate and operator in mean projection ratings be predicted from lexical semantic/pragmatic properties of the predicates, and, if so, how?
	This is a pressing question for future research, to which our data offer some tentative answers. We identify four major patterns.
	%
	The predicates \emph{pretend} and \emph{think} exhibit the {\bf `Negation high'} pattern, shown in panel (a) of \textbf{Figure~2}: We tentatively hypothesize that negation (but not the other operators) interacts with the semantic or pragmatic antiveridicality contributed by these predicates.
	%
	The inferential predicates \emph{prove, confirm}, and \emph{establish} exhibit a {\bf Negation low'} pattern, shown in (b): Here, we tentatively hypothesize that the veridical meaning component interacts with negation (but not the other operators), to result in lower projection ratings under negation.
	%
	For \emph{announce, confess, admit}, and \emph{reveal}, the CC is most projective when embedded in the antecedent of a conditional: This {\bf `Conditional high'} pattern, shown in panel (c), may suggest that the discourse effect of a conditional interacts with the change-of-state communication predicates.
	%
	Finally, the predicates  \emph{inform, know}, and \emph{be annoyed} exhibit a {\bf `Modal low'} pattern, shown in (d). The lexical meaning of these predicates, whose CCs are among the most projective, appears to interact with the modal adverb {\em perhaps}, yielding lower projection ratings.

@@

\newpage
{\scriptsize
\noindent {\bf (Selected) References:}
%
{\bf Beaver (2010).} Have you noticed that your belly button lint colour is related to the colour of your clothing? {\em Presuppositions and Discourse: Essays offered to Hans Kamp.}~\textbullet~
{\bf Degen \& Tonhauser (2022).} Are there factive predicates? An empirical investigation. {\em Language.}~\textbullet~
{\bf Djärv \& Bacovcin (2017).} Prosodic effects on factive presupposition projection. {\em Semantics and Linguistic Theory.}~\textbullet~
{\bf Djärv, Zehr \& Schwarz (2018).} Cognitive vs. emotive factives: An experimental differentiation. {\em Proceedings of Sinn und Bedeutung.}~\textbullet~
{\bf Hooper \& Thompson (1973).} On the applicability of root transformations. {\em Linguistic Inquiry.} ~\textbullet~
{\bf Karttunen (1971).} Some observations on factivity. {\em Research on Language \& Social Interaction.} ~\textbullet~
{\bf Mahler (2020).} The social component of projection behavior of clausal complements. {\em Linguistic Society of America.}~\textbullet~
{\bf Smith and Hall (2014).} The relationship between projection and embedding environment. {\em Proceedings of the 48th Meeting of the Chicago Linguistics Society.}~\textbullet~
{\bf Tonhauser, Beaver, \& Degen (2018).} How projective is projective content? Gradience in projectivity and at-issueness. {\em Journal of Semantics.}~\textbullet~
{\bf Tönnis (2021).} {\em German es-clefts in discourse. A question-based analysis involving expectedness.} Doctoral dissertation, Graz University.

\begin{figure}[h]
	\vspace{-.8\baselineskip}
	\centering
	\scalebox{.9}{\includegraphics[width=\textwidth]{graphs/proj-by-both.pdf}}\vspace{-1.7\baselineskip}
	\caption{\small Mean certainty ratings by predicate and operator with 95\% bootstrapped confidence intervals. Embedding operator coded by letter and color:  \texttt{N} (orange): negation, \texttt{M} (black): modals, \texttt{C} (green): conditional antecedents, \texttt{Q} (blue): polar questions.}
	\label{fig:figure1}
\end{figure}

>>>>

\vspace{-.6\baselineskip}
\noindent 
	We present experimental evidence \textbf{(i)} that the projection of the content of attitude complements differs between entailment-canceling operators, and \textbf{(ii)} that this by-operator variation differs between attitude predicates. Our results do not align with the long-standing distinction of factive vs. semi-factive predicates (see, e.g., \citealp{karttunen_observations_1971,hooper_applicability_1973,djarv_cognitive_2018}).
	Instead, the observed by-operator variation reveals a more nuanced picture of lexical semantic/pragma-tic properties which raises important questions for future research on projection.

@@

\vspace{-.4\baselineskip}
	\citet{karttunen_observations_1971} proposed that the CC of factive predicates (e.g. \emph{regret, forget}) projects across all four operators, whereas that of semi-factive predicates (e.g. \emph{discover, realize, see, notice}) always projects from under negation, but not always from under polar questions, modals, or conditionals. 
	To date, systematic empirical investigations of Karttunen's intuition are scarce.

@@

\noindent
{\bf Discussion: Novel research question.}
	Can the observed interaction between predicate and operator in mean projection ratings be predicted from lexical semantic/pragmatic properties of the predicates, and, if so, how?
	This is a pressing question for future research, to which our data offer some tentative answers. We identify four major patterns.
	%
	The predicates \emph{pretend} and \emph{think} exhibit the {\bf `Negation high'} pattern, shown in panel (a) of \textbf{Figure~2}: We tentatively hypothesize that negation (but not the other operators) interacts with the semantic or pragmatic antiveridicality contributed by these predicates.
	%
	The inferential predicates \emph{prove, confirm}, and \emph{establish} exhibit a {\bf Negation low'} pattern, shown in panel (b): Here, we tentatively hypothesize that the veridical meaning component interacts with negation (but not the other operators), to result in lower projection ratings under negation.
	%
	For \emph{announce, confess, admit}, and \emph{reveal}, the CC is most projective when embedded in the antecedent of a conditional: This {\bf `Conditional high'} pattern, shown in (c), may suggest that the discourse effect of a conditional interacts with the change-of-state communication predicates.
	%
	Finally, the predicates  \emph{inform, know}, and \emph{be annoyed} exhibit a {\bf `Modal low'} pattern, shown in (d). The lexical meaning of these predicates, whose CCs are among the most projective, appears to interact with the modal adverb {\em perhaps}, yielding lower projection ratings.

@@

\newpage
{\scriptsize
\noindent {\bf References:}
%
{\bf Beaver (2010).} Have you noticed that your belly button lint colour is related to the colour of your clothing? {\em Presuppositions and Discourse: Essays offered to Hans Kamp.}~\textbullet~
{\bf Degen \& Tonhauser (2022).} Are there factive predicates? An empirical investigation. {\em Language.}~\textbullet~
{\bf Djärv \& Bacovcin (2017).} Prosodic effects on factive presupposition projection. {\em Semantics and Linguistic Theory.}~\textbullet~
{\bf Djärv, Zehr \& Schwarz (2018).} Cognitive vs. emotive factives: An experimental differentiation. {\em Proceedings of Sinn und Bedeutung.}~\textbullet~
{\bf Hooper \& Thompson (1973).} On the applicability of root transformations. {\em Linguistic Inquiry.} ~\textbullet~
{\bf Karttunen (1971).} Some observations on factivity. {\em Research on Language \& Social Interaction.} ~\textbullet~
{\bf Mahler (2020).} The social component of projection behavior of clausal complements. {\em Linguistic Society of America.}~\textbullet~
{\bf Smith and Hall (2014).} The relationship between projection and embedding environment. {\em Proceedings of the 48th Meeting of the Chicago Linguistics Society.}~\textbullet~
{\bf Tonhauser, Beaver, \& Degen (2018).} How projective is projective content? Gradience in projectivity and at-issueness. {\em Journal of Semantics.}~\textbullet~
{\bf Tönnis (2021).} {\em German es-clefts in discourse. A question-based analysis involving expectedness.} Doctoral dissertation, Graz University.
{\bf Zimmermann (2011).} {\em The grammatical expression of focus in West Chadic: Variation and uniformity in and across languages.} Walter de Gruyter.

\begin{figure}[h]
	\vspace{-.6\baselineskip}
	\centering
	\scalebox{.9}{\includegraphics[width=\textwidth]{graphs/proj-by-both.pdf}}\vspace{-1.7\baselineskip}
	\caption{\small Mean certainty ratings by predicate and operator with 95\% bootstrapped confidence intervals. Embedding operator coded by letter and color:  \texttt{N} (orange): negation, \texttt{M} (black): modals, \texttt{C} (green): conditional antecedents, \texttt{Q} (blue): polar questions.}
	\label{fig:figure1}
\end{figure}
